Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-11-07 doc/paper-lcr.tex

\Fref{fig:where} shows the cumulative contribution to $\kappah$ as a function
of projected distance from the line of sight, magnitude and redshift.
We find that most of the convergence comes from halos whose
galaxies are close to the line of sight. 
Over half of the  convergence
typically comes from halos within 30 arcseconds of the line of sight; by 2
arcminutes, this fraction is 85\%. The contribution from halos beyond 2 arcminutes is relatively constant with a contribution of 0.008$\pm$0.003. We find that ignoring halos beyond 2 arcminutes has no effect on the precision of the reconstruction; this is shown in Figure \ref{fig:radcut}.
As a function of magnitude, we find
that $\kappah$ is dominated by objects with magnitudes between $i=18$ and
$i=24$. Objects brighter than $i=18$ are either too rare, or too close to the
observer, to make a significant contribution to the convergence. Objects
fainter than $i=24$ are too small to be important, unless they are extremely
close to the line of sight, but we find that including halos fainter than
$i=24$ does not improve the reconstruction.  \Fref{fig:magcut} shows the
scatter on $\kappah-\kappa$ (where $\kappah$ is shifted so that its mean is
zero) as a function of reconstruction magnitude limit.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figs/radius_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappa$ as a function of the limiting radius of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:magcut}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/biasplots.eps}
\caption{Estimating $\kappa-\kappa_{\rm true}$ from subsets of
lenses. By subtracting the true (ray-traced) convergence from the
infered P($\kappa$) for each line of sight we are left with a PDF that
should be estimating zero, these can be multiplied together for multiple
lines of sight to test for a possible bias. These plots show the
expectation value of $\prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})$ --
deviations from zero constitute biases. The solid, dashed and dotted
lines correspond to combinations of 
20, 5 and 2 sightlines respectively. Green lines show the results
from multiplying $\pr(\kappa|\data)$ \infered from photometry alone, whilst red
lines are from multiplying $\pr(\kappa)$ \infered from photometry and a
spectroscopic redshift for all halos with $i<23$ and all halos within 1
arcminute and $i<24$. The leftmost plot shows randomly selected lines
of sight. The central plot shows sight lines randomly selected from the
33\% of lines whose P($\kappa$) is most tightly constrained. The right
hand plot shows only lines of sight with external shear of 0.03 or
greater.}
\label{fig:biasplots}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{}
\label{fig:no-median-tricks}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The reduction of $\pr(\kappah|\data)$ to just its median value, before using
it to look up the appropriate $\pr(\kappa)$ from the simulated sightline
ensemble, is reminiscent of the procedure followed by \citet{SuyuEtal2010},
and Greene et al (in prep.) The reconstruction method can be seen as the
limiting case of the weightings explored by Greene et al: rather than using
weighted number counts in an aperture, we treat each galaxy individually, and
predict  $\kappa_{\rm h,obs}^{\rm med}$ for use in their place. One
significant disadvantage of these types of method is that they impose a
bottleneck in the flow of information. Once you have compressed to one number,
it is difficult to incorporate more information later; you have to recompute
all the calibrations. Using the full PDF for $\kappah$ is the more correct
thing to do: at present, however, it suffers from significant bias in the
photometric data case. \Fref{fig:no-median-tricks} shows how $\kappa$ is
overestimated, as a result of the high scatter in halo mass for objects with 
high stellar mass. As already noted, 
future work should increasing the precision of the halo
mass estimates.

>>>>

\Fref{fig:where} shows the cumulative contribution to $\kappah$ as a function
of projected distance from the line of sight, magnitude and redshift. We find
that most of the convergence comes from halos whose galaxies are close to the
line of sight.  Over half of the  convergence typically comes from halos
within 30 arcseconds of the line of sight; by 2 arcminutes, this fraction is
85\%. The contribution from halos beyond 2 arcminutes is relatively constant
at a contribution of 0.008$\pm$0.003. We find that ignoring halos beyond 2
arcminutes has no effect on the precision of the reconstruction; this is shown
in \Fref{fig:radcut}. As a function of magnitude, we find that $\kappah$
is dominated by objects with magnitudes between $i=18$ and $i=24$. Objects
brighter than $i=18$ are either too rare, or too close to the observer, to
make a significant contribution to the convergence. Objects fainter than
$i=24$ are too small to be important, unless they are extremely close to the
line of sight, but we find that including halos fainter than $i=24$ does not
improve the reconstruction.  \Fref{fig:magcut} shows the scatter on
$\kappah-\kappa$ (where $\kappah$ is shifted so that its mean is zero) as a
function of reconstruction magnitude limit.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figs/radius_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappa$ as a function of the limiting radius of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:radcut}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/biasplots.eps}
\caption{Convergence accuracy from combining subsets of
lenses. By subtracting the true (ray-traced) convergence from the
\infered $\pr(\kappa)$ for each line of sight we are left with a PDF that
should be estimating zero; these can be multiplied together for multiple
lines of sight to emulate a joint likelihood analysis, and 
test for possible bias in it. These plots show the
expectation value of $\prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})$ --
deviations from zero constitute biases. The solid, dashed and dotted
lines correspond to combinations of 
20, 5 and 2 sightlines respectively. Green lines show the results
from multiplying $\pr(\kappa|\data)$ \infered from photometry alone, whilst red
lines are from multiplying $\pr(\kappa)$ \infered from photometry and a
spectroscopic redshift for all halos with $i<23$ and all halos within 1
arcminute and $i<24$. The leftmost plot shows randomly selected lines
of sight. The central plot shows sight lines randomly selected from the
33\% of lines whose P($\kappa$) is most tightly constrained. The right
hand plot shows only lines of sight with external shear of 0.03 or
greater.}
\label{fig:biasplots}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{Convergence estimation accuracy using the full  $\pr(\kappa|\kappah)$
distribution, rather than its median.  As in \Fref{fig:biasplots}, solid,
dashed and dotted lines show combinations of 20, 5 and 2 lines of sight
respectively, while green lines show the inferences 
from photometry alone, red lines from photometry and targeted spectroscopy 
(assuming a spectroscopic redshift for all
halos with $i<23$ and all halos within 1 arcminute and $i<24$), and blue lines
from the unrealistic assumption of a spectroscopic redshift for all galaxies
in the field. }
\label{fig:no-median-tricks}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The reduction of $\pr(\kappah|\data)$ to just its median value, before using
it to look up the appropriate $\pr(\kappa)$ from the simulated sightline
ensemble, is reminiscent of the procedure followed by \citet{SuyuEtal2010},
and Greene et al (in prep.) The reconstruction method can be seen as the
limiting case of the weightings explored by Greene et al: rather than using
weighted number counts in an aperture, we treat each galaxy individually, and
predict  $\kappa_{\rm h,obs}^{\rm med}$ for use in their place. One
significant disadvantage of these types of method is that they impose a
bottleneck in the flow of information. Once you have compressed to one number,
it is difficult to incorporate more information later; you have to recompute
all the calibrations. Using the full PDF for $\kappah$ is the more correct
thing to do: at present, however, it suffers from significant bias in the
photometric data case. \Fref{fig:no-median-tricks} shows how $\kappa$ is
overestimated, as a result of the high scatter in halo mass for objects with 
high stellar mass. As has already been noted, 
future work should increasing the precision of the halo
mass estimates.
